%% ============================================================================
\section{Score Functions}
\label{app:scores}
%% ============================================================================

For reference, we provide the score functions under the misspecified and true
models for series systems with general component hazards.

\paragraph{Misspecified score (C1-C2-C3).}
The score under the standard model for parameter $\theta_j$ is:
\begin{equation}
    \frac{\partial \ell^{\mathrm{wrong}}}{\partial \theta_j} =
    \sum_{i=1}^n \frac{\partial \log R_j(s_i; \theta_j)}{\partial \theta_j}
    + \sum_{i: \delta_i = 1}
    \frac{\ind{j \in c_i}\, h_j'(s_i; \theta_j)}{\sum_{k \in c_i} h_k(s_i; \theta_k)},
\end{equation}
where $h_j' = \partial h_j / \partial \theta_j$.

\paragraph{True score (C1-C3 with known $\mathbf{P}$).}
Under the Bernoulli masking model with masking weights $\pi_{k,c}$:
\begin{equation}
    \frac{\partial \ell^{\mathrm{true}}}{\partial \theta_j} =
    \sum_{i=1}^n \frac{\partial \log R_j(s_i; \theta_j)}{\partial \theta_j}
    + \sum_{i: \delta_i = 1}
    \frac{\pi_{j,c_i}\,\ind{j \in c_i}\, h_j'(s_i; \theta_j)}
         {\sum_{k \in c_i} h_k(s_i; \theta_k)\,\pi_{k,c_i}}.
\end{equation}

\paragraph{Exponential specialization.}
For exponential components ($h_j = \theta_j$, $R_j = e^{-\theta_j t}$), these
simplify to:
\begin{align}
    \frac{\partial \ell^{\mathrm{wrong}}}{\partial \theta_j}
    &= -\sum_{i=1}^n s_i + \sum_{i: \delta_i = 1}
        \frac{\ind{j \in c_i}}{\sum_{k \in c_i} \theta_k}, \\
    \frac{\partial \ell^{\mathrm{true}}}{\partial \theta_j}
    &= -\sum_{i=1}^n s_i + \sum_{i: \delta_i = 1}
        \frac{\pi_{j,c_i}\,\ind{j \in c_i}}
             {\sum_{k \in c_i} \theta_k\,\pi_{k,c_i}}.
\end{align}

%% ============================================================================
\section{Non-Identifiability Evidence}
\label{sec:appendix-ident}
%% ============================================================================

To demonstrate the non-identifiability of $(\vtheta, \mathbf{P})$, we
attempted joint estimation with $n = 2000$ observations from a 2-component
exponential system with $\vtheta^* = (1.0, 2.0)$ and asymmetric
$\mathbf{P}$ ($p_{21} = 0.10$, $p_{12} = 0.90$).

\begin{center}
\begin{tabular}{lcc}
\toprule
Method & $\hat\theta_1$ & $\hat\theta_2$ \\
\midrule
Joint $(\vtheta, \mathbf{P})$ & 1.62 & 1.44 \\
Known $\mathbf{P}$ & 0.97 & 2.08 \\
True values & 1.00 & 2.00 \\
\bottomrule
\end{tabular}
\end{center}

The joint estimator consistently converges to
$\hat\vtheta \approx (1.6, 1.4)$ with
$\hat{P}_{21} \approx 0.45$ (true: 0.10), regardless of sample size. The
total hazard is well-identified: $\sum_j \hat\theta_j = 3.01$ vs.\ true
3.00. This confirms \Cref{thm:non-ident}: individual rates and masking
probabilities are confounded, but the total system hazard is identifiable.

%% ============================================================================
\section{Set-Valued Masking versus Point-Cause Misclassification}
\label{app:set-vs-point}
%% ============================================================================

A natural question for readers familiar with the biostatistics
competing-risks literature is whether the relaxed-C1 likelihood developed
here can be obtained as a special case of point-cause misclassification with
random masking. Methods for the latter setting, including
\citet{Ebrahimi1996}, \citet{VanRompaye2012}, \citet{HaTsodikov2015}, and
\citet{Bakoyannis2025}, might then apply directly. We show that the two
model families are non-nested and that set-valued masking under C1 violation
cannot, in general, be obtained from point-cause misclassification composed
with random coarsening.

\begin{definition}[Point-cause misclassification with random masking]
\label{def:point-misclass}
The data-generating process is:
\begin{enumerate}
    \item The true failed component $K \in \{1, \ldots, m\}$ is drawn from
        the system-failure distribution.
    \item A recorded cause $K^{\mathrm{obs}} \in \{1, \ldots, m\}$ is drawn
        from a misclassification distribution
        $\pi_{j \mid k} = \Prob\{K^{\mathrm{obs}} = j \mid K = k\}$.
    \item The candidate set $C \subseteq \{1, \ldots, m\}$ is drawn
        conditional on $K^{\mathrm{obs}} = j$ from some distribution
        $\rho(c \mid j)$ satisfying $\rho(c \mid j) > 0$ only if $j \in c$.
\end{enumerate}
\end{definition}

\begin{lemma}[Non-nesting of set-valued masking and point-cause models]
\label{lem:non-nesting}
Let $\mathcal{P}_{\mathrm{set}}(m)$ denote the family of joint distributions
of $(K, C)$ generated by Bernoulli set-valued masking under relaxed C1, in
which conditional on $K = k$ the candidate set $C$ is formed by including
each component $j$ independently with probability $q_j(k) \in [0, 1]$. Let
$\mathcal{P}_{\mathrm{point}}(m)$ denote the family generated by
\Cref{def:point-misclass}. Then for $m \geq 2$,
$\mathcal{P}_{\mathrm{set}}(m)$ and $\mathcal{P}_{\mathrm{point}}(m)$ are
non-nested: neither family is a subset of the other.
\end{lemma}

\begin{proof}
$\mathcal{P}_{\mathrm{set}} \not\subseteq \mathcal{P}_{\mathrm{point}}$:
Choose $q_j(k) \in (0, 1)$ for all $j, k$. Then
\begin{equation*}
    \Prob\{C = \emptyset \mid K = k\} = \prod_{j=1}^{m} (1 - q_j(k)) > 0.
\end{equation*}
Under $\mathcal{P}_{\mathrm{point}}$, the candidate set $C$ contains
$K^{\mathrm{obs}}$ almost surely, so $C \neq \emptyset$ almost surely.

$\mathcal{P}_{\mathrm{point}} \not\subseteq \mathcal{P}_{\mathrm{set}}$: For
$m = 4$ and any $K$, take $\pi_{2 \mid 1} = \pi_{3 \mid 1} = 1/2$,
$\rho(\{2, 4\} \mid 2) = 1$, $\rho(\{3, 4\} \mid 3) = 1$. Then conditional on
$K = 1$:
\begin{align*}
    \Prob\{C = \{2, 4\} \mid K = 1\} &= 1/2, \\
    \Prob\{C = \{3, 4\} \mid K = 1\} &= 1/2, \\
    \Prob\{2 \in C, \, 3 \in C \mid K = 1\} &= 0.
\end{align*}
Under $\mathcal{P}_{\mathrm{set}}$ with marginal inclusion probabilities
$q_2(1) = \Prob\{2 \in C \mid K = 1\} = 1/2$ and $q_3(1) = 1/2$, independence
of inclusion implies
\begin{equation*}
    \Prob\{2 \in C, \, 3 \in C \mid K = 1\}
        = q_2(1)\, q_3(1) = 1/4 \neq 0,
\end{equation*}
contradicting the joint distribution above.
\end{proof}

\begin{remark}[Consequences for inference]
\label{rem:set-vs-point}
Direction (i) of \Cref{lem:non-nesting} shows that relaxed-C1 distributions
cannot be recovered from point-cause misclassification models without
explicitly admitting an ``empty candidate set'' outcome that is not part of
those models' standard formulations. Direction (ii) shows that point-cause
misclassification with arbitrary masking conditional on the recorded cause
can produce correlation structures in candidate-set inclusion that the
Bernoulli set-valued model cannot reproduce.

Practically, this means that estimators and sensitivity-analysis tools
developed for cause-of-failure misclassification, in particular those of
\citet{Ebrahimi1996}, \citet{VanRompaye2012}, \citet{HaTsodikov2015},
\citet{MorenoBetancur2015}, and \citet{Bakoyannis2025}, are not directly
applicable to the set-valued candidate-set setting studied in this paper.
Conversely, the relaxed-C1 analysis here does not subsume those methods for
the point-cause setting. The two literatures address related but distinct
observational structures.
\end{remark}

%% ============================================================================
\section{Software}
\label{app:software}
%% ============================================================================

All methods are implemented in the \texttt{mdrelax} R package (version 1.0.0),
available at \url{https://github.com/queelius/mdrelax}. The package provides:
\begin{itemize}
    \item Data generation under C1-C2-C3 and relaxed C2 for both
        exponential and Weibull series systems.
    \item MLE with analytical score and Fisher information for all model tiers.
    \item The $\mathbf{P}$ matrix construction and masking probability
        computation functions used in this paper.
    \item 1276 unit tests verifying likelihood correctness, score--gradient
        agreement, FIM consistency, and MLE convergence properties.
\end{itemize}

The sensitivity sweep (\texttt{paper/run\_sensitivity\_sweep.R}) reproduces
all figures and tables in \Cref{sec:simulations}.
