%% ============================================================================
\section{Conclusion}
\label{sec:conclusion}
%% ============================================================================

We have studied the sensitivity of maximum-likelihood estimation in masked
series systems to violations of the three coarsening conditions C1, C2,
and C3. The contributions of the paper are unified by the robustness
hierarchy that emerges from the analysis:

\begin{enumerate}
    \item \textbf{Calibrated severity scales for each of the three
        coarsening violations.} Each violation is parameterized on a
        scalar severity $\alpha_{C_k} \in [0, \alpha_{C_k}^{\max}]$ with
        $\alpha_{C_k} = 0$ recovering the standard model and
        $\alpha_{C_k}^{\max}$ marking the boundary to adversarial
        regimes. Together they form a multi-dimensional sensitivity
        geometry that complements the calibrated-robustness-value
        philosophy of \citet{CinelliHazlett2020} and the pattern-mixture
        sensitivity parameter of \citet{MorenoBetancur2015}.

    \item \textbf{Closed-form local sensitivity indices specializing
        ISNI to set-valued masking.} The first-order expansions of the
        misspecified MLE at the standard-model fit are derived in
        closed form for each violation. For exponential components the
        system-hazard preservation is exact, not first-order, by a
        profile-likelihood factorization argument
        (\Cref{thm:exact-hazard-exp}); for Weibull components the
        expansion specializes the ISNI construction of
        \citet{TroxelMaHeitjan2004} and \citet{ZhangHeitjan2006} from
        interval-censoring to candidate-set coarsening.

    \item \textbf{Robustness intervals for practitioner-facing
        estimands.} For each violation type and each reliability
        functional of interest, the robustness interval gives the range
        of severities within which the estimate remains within a stated
        tolerance of the standard-model fit. The implementation
        (Monte-Carlo and first-order variants) is available in the
        \texttt{mdrelax} R package.

    \item \textbf{Robustness hierarchy classifying estimands.}
        System-level estimands (total hazard, MTTF, system reliability)
        are robust to all three violations: exactly for exponential
        components, at first order for Weibull under C2 and C3, and
        with a small bounded first-order bias under C1 for Weibull.
        Component-level estimands are fragile under all three
        violations regardless of component family. The simulation
        sweeps validate the hierarchy at two sample sizes with
        identical qualitative conclusions.

    \item \textbf{Reproducible software.} The methods are implemented
        in the \texttt{mdrelax} R package, which provides data
        generation, MLE, Fisher information, and robustness-interval
        computation for all model tiers. None of the closest competing
        methodology papers release code.
\end{enumerate}

The headline message for the practitioner is sharp. A reliability
engineer reporting system-level estimates from the standard
masked-data Weibull MLE has substantial protection against any
plausible violation of the masking-condition assumptions. A
reliability engineer reporting component-level estimates does not,
and should perform sensitivity analysis using the methods of
\Cref{sec:hierarchy} before publication.

All code, simulations, and the manuscript source are available at
\url{https://github.com/queelius/mdrelax}. The k-out-of-m generalization
of the masking framework, including a threshold theorem characterizing
when individual component parameters are identifiable, is developed in
companion work \citep{towell2026binary}.
