%% ============================================================================
\section{Discussion}
\label{sec:discussion}
%% ============================================================================

The theoretical results of \Cref{sec:framework} and the empirical sweeps of
\Cref{sec:simulations} together establish a robustness hierarchy for
masked series-system reliability inference. We summarize the practical
implications, structural reasons behind the hierarchy, limitations of the
analysis, and natural extensions.

\subsection{When Does Each Violation Matter?}
\label{sec:when-matter}

The robustness hierarchy of \Cref{cor:hierarchy} partitions estimands into
three regimes:

\paragraph{System-level estimands are reliable across all three violations.}
For the total system hazard, mean time to failure, and system reliability
at a fixed time, the standard C1-C2-C3 MLE produces estimates whose bias
remains bounded across the full range of practical violation severities.
For exponential components, the preservation is exact under any masking
violation (\Cref{thm:exact-hazard-exp}); the profile-likelihood
factorization decouples the system-rate score from the candidate-set
information entirely. For Weibull components, preservation is first-order
under C2 and C3 violations and admits a small bounded first-order bias
under C1 violations. The empirical bound at the 5-component system
studied here is under 4\% relative bias at $\alpha_{C_1} = 0.5$, and
under 1\% under C2 or C3.

\paragraph{Component-level estimands are fragile under any violation.}
Individual component parameters (rates for exponential, shape-scale pairs
for Weibull) are not first-order robust to any of the three violations.
The empirical results document the failure mode in stark detail:
component-level relative biases can exceed 100\% (e.g., $\lambda_5$ under
C3 violation), confidence-interval coverage can collapse from nominal to
zero (e.g., $\lambda_1$ under C3), and the Fisher information for
specific parameter combinations can become near-singular under C2
perturbations (the $\lambda_2$ instability at $s = 0.5$). The standard
MLE makes confident but systematically wrong claims about which
components are unreliable.

\paragraph{The asymmetry between C1 and the other two violations for
Weibull.} The simulation sweep cleanly separates the three violations by
their first-order behavior. Under C2 and C3, the linear coefficient
in the pseudo-true expansion is statistically indistinguishable from
zero at $p > 0.20$ at both $n = 500$ and $n = 2000$, confirming
first-order preservation. Under C1, the linear coefficient is consistent
and significantly nonzero ($\sim -0.30$ at both sample sizes,
$p < 0.001$). The structural reason traces to where each violation
inserts bias into the likelihood: C2 and C3 distort only the
candidate-set portion (which couples to the time portion at second order
through the Weibull shape-scale interaction), while C1 violation
introduces empty or wrong-cause candidate sets that bias the
candidate-set portion more aggressively, and that bias propagates
into the time-portion estimates at first order through the shared
parameterization.

\subsection{Practical Guidance}
\label{sec:guidance}

Since the masking mechanism is a diagnostic process rarely known in
detail, the practitioner usually cannot verify whether C1, C2, or C3
holds. The robustness hierarchy yields a simple decision procedure.

\begin{enumerate}
    \item \textbf{Fit the standard model.} The C1-C2-C3 MLE is
        well-understood, computationally straightforward, and yields
        reliable system-level estimates regardless of coarsening
        condition status under the regimes characterized in this paper.

    \item \textbf{Identify the inferential target.} If only system-level
        quantities are needed (system hazard, MTTF, system survival at a
        fixed time), the standard model suffices. No correction or
        sensitivity analysis is required for exponential components, and
        only mild caution is warranted for Weibull components under
        suspected C1 violation.

    \item \textbf{For component-level inference, perform sensitivity
        analysis.} Construct a family of plausible masking models
        representing different violation directions and severities.
        Refit the standard model under each scenario or use the local
        sensitivity indices of \Cref{sec:hierarchy} to bound the
        first-order bias. If the component-level conclusions are stable
        across the family, the violations are not a concern for the
        application at hand.

    \item \textbf{Report sensitivity bounds, not point estimates, when
        conclusions are violation-sensitive.} For component-level claims
        that move under the assumed masking structure, report the range
        of estimates as a sensitivity bound. This parallels the partial
        identification approach \citep{manski2003partial}: data identify
        a set of possible component parameter values rather than a
        point estimate, and reporting the set is more honest than
        reporting a single estimate under an unverifiable assumption.
\end{enumerate}

\subsection{Limitations}
\label{sec:limitations}

Several aspects of the analysis warrant explicit caveats.

\begin{itemize}
    \item \textbf{Parametric perturbation families.} The Bernoulli C1,
        directional-P C2, and power-weight C3 sweeps are specific
        parametric violations. Each was chosen to be interpretable and
        cover a calibrated severity scale, but other violation
        structures (e.g., masking that depends explicitly on $t$,
        diagnostic procedures that interact with system age, masking
        that depends on observed covariates) are not covered. The
        first-order preservation arguments of \Cref{sec:misspec}
        and \Cref{sec:c3-sensitivity} extend in principle to any
        smooth violation family, but the empirical magnitudes depend on
        the specific perturbation.

    \item \textbf{Lifetime distributions.} We study Weibull components,
        with exponential as a constant-hazard special case. Other
        parametric families (log-normal, Gompertz, gamma, generalized
        gamma) have different hazard shapes and may exhibit different
        first-order behavior under the same violations. The exponential
        exact-preservation result (\Cref{thm:exact-hazard-exp}) depends
        specifically on the additive factorization
        $\ell(S, \phi) = \ell_S(S) + \ell_\phi(\phi)$ that does not
        generalize to non-constant-hazard models.

    \item \textbf{Direction dependence for C2.} The specific direction
        matrix $\mathbf{D}$ in the C2 perturbation determines which
        component parameters absorb the masking asymmetry. Different
        $\mathbf{D}$ matrices produce different component-level bias
        patterns. The system-hazard robustness persists by
        \Cref{prop:hazard-robust} regardless of $\mathbf{D}$, but
        component-level fragility claims should be interpreted as
        representative rather than universal.

    \item \textbf{Sample size.} The main simulations use $n = 500$
        with $n = 2000$ as a robustness check. The bias trajectories
        are qualitatively stable across the $4\times$ sample-size
        change, but coverage degradation is more pronounced at larger
        $n$ (confidence intervals narrow while bias persists), an
        effect already visible in the $\lambda_1$ coverage trajectory
        under C3.

    \item \textbf{Variance estimation.} Coverage uses Wald intervals
        based on model-based Fisher information, which assumes correct
        specification. Under masking-condition violations, the sandwich
        estimator \citep{huber1967behavior, white1982misspecified} is
        more appropriate. The observed coverage degradation conflates
        bias and incorrect standard errors; disentangling them via
        sandwich-based coverage is straightforward to add and is left
        for the software extension.

    \item \textbf{Set-valued versus point-cause masking.} The framework
        and results apply to set-valued candidate-set masking. As
        \Cref{app:set-vs-point} shows, this observational structure
        does not reduce to point-cause misclassification with random
        coarsening, so the methodology of \citet{Ebrahimi1996},
        \citet{VanRompaye2012}, \citet{MorenoBetancur2015}, and
        \citet{Bakoyannis2025} is not directly applicable.
\end{itemize}

\subsection{Future Work}
\label{sec:future}

Three natural extensions are visible from the current results.

\paragraph{A formal specification test for $\{C_1, C_2, C_3\}$.}
Under $H_0 \colon \{C_1, C_2, C_3\}$, the conditional distribution of
the candidate set given the failure time factors as
$\Pr(C = c \mid T = t) = \pi_c \cdot R_c(t; \vtheta^*)$ where
$\pi_c$ depends only on $c$ and $R_c(t; \vtheta) = \sum_{k \in c}
h_k(t; \theta_k) / S(t; \vtheta)$ is the relative-hazard contribution.
This factorization is testable from observed data alone (no need to
observe the failed cause $K$). A bin-based Pearson statistic on
candidate-set frequencies provides a workable null distribution; the
practical challenge is power against the C1 violation (the most
inferentially consequential) at typical reliability-study sample
sizes. We have implemented and validated the singleton version of the
test in our companion software; substantial development is needed for
a publication-ready version, including the all-candidate-set extension
and a closed-form sandwich variance.

\paragraph{Second-order theory for the C1 first-order failure.}
The simulation results show a quadratic-in-$\alpha$ trajectory for the
Weibull C1 bias, with a significant quadratic curvature at both
sample sizes. Deriving the second-order coefficient
$\Delta^{(2)}_j(\vtheta^*; p_0)$ would explain the bias-bounding
phenomenon (why the bias plateaus rather than growing linearly) and
provide tighter robustness intervals than the first-order ISNI alone
can deliver. The derivation should follow from extending the proof of
\Cref{thm:misspec-C1} to the second-order expansion of the score
equations.

\paragraph{Extension to other parametric families.}
The exponential-versus-Weibull asymmetry (exact preservation versus
first-order preservation) follows from the constant-hazard property of
the exponential. Parametric families with related structural properties
(gamma at integer shape, log-logistic with specific shape constraints,
generalized exponential subfamilies) may admit analogous exact
preservation results. A systematic survey of which masked-data
parametric families admit which preservation tier would strengthen
the framework's generality.
