%% ============================================================================
\section{Introduction}
\label{sec:introduction}
%% ============================================================================

Estimating the reliability of individual components in a series system is a
fundamental challenge when failure causes are masked. When a series system
fails, diagnostic procedures may identify only a \emph{candidate set} of
components that could have caused the failure, rather than pinpointing the
exact failed component \citep{Usher-1988}. Maximum likelihood estimation under
masked data has been developed extensively under a standard set of
assumptions~\citep{Lin-1993, towell2023reliability, towell2026masked}: (C1)~the true failed
component is always in the candidate set; (C2)~masking is non-informative,
meaning the candidate set formation is independent of the failure cause given
the failure time; and (C3)~the masking mechanism does not depend on the
component reliability parameters.

Each of these conditions is a restriction on the diagnostic process that
produces the candidate set. Taken together, they are equivalent to the joint
assumption of \emph{coarsening at random}
\citep{HeitjanRubin1991, JacobsenKeiding1995} and parameter distinctness
\citep{Rubin1976} for the candidate-set coarsening map. Each may fail in
practice. The technician's experience may make the candidate set informative
about the cause (C2 violation). An automated diagnostic may weight candidate
inclusion by failure-rate ranking (C3 violation). In the worst case, the
diagnostic may miss the failed component entirely, producing a candidate set
that does not contain the true cause (C1 violation).

Prior work has addressed each violation in isolation, often in distinct
vocabularies. For C2, alternative parametric masking models have been
proposed in the reliability literature \citep{LinGuess1994, Guttman1995,
Mukhopadhyay2006, CraiuReiser2006}. The analogous problem of cause
uncertainty in biostatistics has produced parallel methodology:
\citet{MorenoBetancur2015} developed pattern-mixture sensitivity analysis
for competing risks regression with missing causes;
\citet{BakoyannisYiannoutsos2015} derived closed-form bias expressions and
permissible-range bounds for cumulative incidence under cause
misclassification; \citet{Bakoyannis2025}, in concurrent work, introduced
robustness intervals for Cox cause-specific hazards under binary missing
cause. The local sensitivity of an MLE to assumption violations has been
formalized as the Index of Local Sensitivity to Nonignorability (ISNI)
\citep{TroxelMaHeitjan2004, ZhangHeitjan2006}, with closely related
first-order bias expansions in \citet{CopasEguchi2005} and
\citet{SiannisCopasLu2005}, and the modern calibrated-bias-bound tradition
culminating in the omitted-variable robustness value of
\citet{CinelliHazlett2020}.

Three structural features of the masked series-system problem are not
addressed in this corpus. First, the candidate set is \emph{set-valued}, not
a single label that may be wrong. Every prior treatment of cause uncertainty
in competing risks treats the cause as either fully observed, fully missing,
or misclassified to a single alternative label; the set-valued case requires
a different likelihood structure and does not reduce to point-cause
misclassification with random masking (\Cref{app:set-vs-point}). Second, the
standard reliability model imposes \emph{three orthogonal coarsening
conditions} (C1, C2, C3) whose violations carry distinct semantics, where
existing sensitivity frameworks operate over a scalar departure parameter.
Third, parametric reliability models (Weibull, exponential series systems)
admit closed-form bias expressions that the predominantly semiparametric
biostatistics literature has not derived. \citet{Bakoyannis2025} explicitly
note that ``there is no formula for computing the E-value in the context of
competing risks analysis with MNAR causes of failure.''

We address these gaps with a unified sensitivity-analysis framework for
masked series-system reliability under calibrated violations of C1, C2, and
C3. Our contributions are:

\begin{enumerate}
    \item \textbf{Calibrated severity scales.} We parameterize each
        coarsening condition on a scalar severity
        $\alpha_{C_k} \in [0, \alpha_{C_k}^{\max}]$ with boundary semantics
        that $\alpha_{C_k} = 0$ recovers the standard C1-C2-C3 model and
        $\alpha_{C_k}^{\max}$ marks the transition to an adversarial regime.
        This aligns with the calibrated-robustness-value philosophy of
        \citet{CinelliHazlett2020} and the pattern-mixture sensitivity
        parameter of \citet{MorenoBetancur2015}, but extends both to a
        multi-dimensional sensitivity geometry.

    \item \textbf{Local sensitivity indices for set-valued masking.} We
        derive closed-form first-order expansions of the misspecified MLE in
        each $\alpha_{C_k}$ at the standard-model fit, specializing the ISNI
        construction of \citet{TroxelMaHeitjan2004} and
        \citet{ZhangHeitjan2006} from interval-censored time to
        candidate-set coarsening. The expansions are explicit for
        exponential components and tractable for Weibull components,
        addressing the gap that \citet{Bakoyannis2025} identify in the
        cumulative-incidence setting.

    \item \textbf{Robustness intervals.} For each violation type and each
        estimand of practical interest (component rates, system hazard,
        mean time to failure), we construct robustness intervals giving the
        range of $\alpha_{C_k}$ within which inferential conclusions
        persist. This extends the framework of \citet{Bakoyannis2025} from
        semiparametric Cox regression with binary missing cause to
        parametric series systems under set-valued candidate-set
        coarsening.

    \item \textbf{Robustness hierarchy across estimands.} We classify which
        functionals of the parameter vector are protected versus fragile
        under each violation type. System-level estimands (total hazard,
        mean time to failure) remain consistently estimated under
        violations of moderate severity, while component-level estimands
        degrade in a violation-specific pattern admitting closed-form
        characterization for exponential systems.

    \item \textbf{Software.} The methods are implemented in the
        \texttt{mdrelax} R package, which provides data generation, MLE,
        Fisher information, sensitivity-index computation, and
        robustness-interval construction for all model tiers. None of the
        closest related work \citep{MorenoBetancur2015,
        BakoyannisYiannoutsos2015, ZhangHeitjan2006, Bakoyannis2025}
        releases code.
\end{enumerate}

The contributions are linked by the joint coarsening structure of C1, C2,
and C3. Together they amount to a sensitivity-analysis toolkit for the
practitioner who has fit the standard masked-series-system MLE and wishes
to know how robust the resulting reliability estimates are to plausible
violations of the modeling assumptions.

We focus throughout on series systems, the $k = 1$ case of the general
$k$-out-of-$m$ family. Companion work \citep{towell2026binary} treats the
complementary identifiability question across general $k$: there the
obstacle is a permutation symmetry of the system likelihood, resolved by a
threshold theorem showing that component-level observations of essentially
any quality break the symmetry. The present paper instead takes the
candidate-set information as given and asks the robustness question: how
much do violations of the masking conditions distort the standard
C1-C2-C3 estimates?

The remainder of this paper is organized as follows.
\Cref{sec:background} reviews series systems, masked data, the standard
C1-C2-C3 likelihood, and the coarsening-at-random framework that subsumes
the three conditions. \Cref{sec:framework} develops the unified sensitivity
framework: calibrated severity scales, the misspecification theorems and
pseudo-true parameters under each violation, local sensitivity indices,
and robustness-interval construction. \Cref{sec:simulations} presents
the simulation sweep across all three violation types.
\Cref{sec:application} demonstrates the workflow on the turbine engine data
of \citet{Huairu-2013}. \Cref{sec:discussion} discusses practical
implications and limitations, and \Cref{sec:conclusion} concludes.
