%% ============================================================================
\section{Sensitivity Framework}
\label{sec:framework}
%% ============================================================================

We now develop the tools needed to study the sensitivity of C1-C2-C3 inference
to C2 violations. We proceed in four steps: the general likelihood under C1
alone (\Cref{sec:like-C1}), a Bernoulli model for generating controlled C2
violations (\Cref{sec:bernoulli}), a misspecification theorem characterizing
the resulting bias (\Cref{sec:misspec}), and a non-identifiability result that
motivates sensitivity analysis over model estimation (\Cref{sec:identifiability}).

\subsection{Likelihood Under C1 Alone}
\label{sec:like-C1}

When only C1 holds, the masking probability enters the likelihood and cannot
be eliminated.

\begin{theorem}[Likelihood Under C1 Alone]
\label{thm:like-C1}
Under Condition C1 alone, the likelihood contribution from an uncensored
observation $(s_i, c_i)$ is:
\begin{equation}
\label{eq:like-C1}
    L_i(\vtheta) = \prod_{\ell=1}^m R_\ell(s_i; \vtheta_\ell) \cdot
    \sum_{k \in c_i} h_k(s_i; \vtheta_k) \cdot
    \Prob\{C_i = c_i \mid T_i = s_i, K_i = k\}.
\end{equation}
\end{theorem}

\begin{proof}
Under C1, $\Prob\{C_i = c \mid T_i = t, K_i = k\} = 0$ when $k \notin c$.
Summing over $K_i$:
\begin{align}
    f_{T_i, C_i}(t, c; \vtheta)
    &= \sum_{k=1}^m h_k(t; \vtheta_k) \prod_{\ell=1}^m R_\ell(t; \vtheta_\ell)
        \cdot \Prob\{C_i = c \mid T_i = t, K_i = k\} \notag \\
    &= \prod_{\ell=1}^m R_\ell(t; \vtheta_\ell) \cdot
        \sum_{k \in c} h_k(t; \vtheta_k)\,
        \Prob\{C_i = c \mid T_i = t, K_i = k\}. \qedhere
\end{align}
\end{proof}

Under C2, the masking probability $\Prob\{C_i = c \mid T_i = t, K_i = k\}$ is
constant over $k \in c$ and factors out of the sum. Under C3, it does not
depend on $\vtheta$ and can be dropped from the likelihood. When either
condition fails, the masking probability remains inside the sum, coupling the
inference about $\vtheta$ to the unknown masking mechanism.

\subsection{The Bernoulli Perturbation Model}
\label{sec:bernoulli}

To generate data with controlled C2 violations, we use a Bernoulli candidate
set model. We do not claim this model describes how masking works in practice.
Rather, it provides a device for producing data whose departure from C2 is
known and calibrated.

\begin{definition}[Bernoulli Masking Model]
\label{def:bernoulli}
Let $p_j(k) = \Prob\{j \in C_i \mid K_i = k\}$ denote the probability that
component $j$ is included in the candidate set given that component $k$
failed. Under C1, $p_j(j) = 1$ for all $j$. These probabilities form an
$m \times m$ matrix:
\begin{equation}
\label{eq:P-matrix}
    \mathbf{P} = \begin{pmatrix}
        1 & p_1(2) & \cdots & p_1(m) \\
        p_2(1) & 1 & \cdots & p_2(m) \\
        \vdots & \vdots & \ddots & \vdots \\
        p_m(1) & p_m(2) & \cdots & 1
    \end{pmatrix},
\end{equation}
where each component is included independently. The probability of observing
candidate set $c$ given $K_i = k$ is
\begin{equation}
\label{eq:pi-k-c}
    \pi_k(c) = \ind{k \in c} \prod_{j \in c \setminus \{k\}} p_j(k)
    \prod_{j \notin c} (1 - p_j(k)).
\end{equation}
\end{definition}

\begin{remark}[C2 in Terms of $\mathbf{P}$]
\label{rem:C2-P}
Condition C2 holds if and only if each row of $\mathbf{P}$ has constant
off-diagonal entries: $p_j(k) = p_j$ for all $k \neq j$. When the columns of
$\mathbf{P}$ differ, the masking mechanism ``knows'' which component failed,
violating C2.
\end{remark}

The likelihood under this model follows immediately from
\Cref{thm:like-C1}:

\begin{corollary}[Likelihood Under the Bernoulli Model]
\label{cor:bernoulli-like}
Under the Bernoulli masking model with known $\mathbf{P}$, the likelihood
contribution from an uncensored observation $(s_i, c_i)$ is:
\begin{equation}
\label{eq:bernoulli-like}
    L_i(\vtheta) = \prod_{\ell=1}^m R_\ell(s_i; \vtheta_\ell) \cdot
    \sum_{k \in c_i} h_k(s_i; \vtheta_k) \cdot \pi_k(c_i).
\end{equation}
\end{corollary}

\paragraph{Parameterizing violation severity.}
We generate a family of $\mathbf{P}$ matrices indexed by a scalar severity
parameter $s \in [0,1]$:
\begin{equation}
\label{eq:P-sweep}
    \mathbf{P}(s) = \mathbf{P}_0 + s\,\mathbf{D},
\end{equation}
where $\mathbf{P}_0$ is the uniform (C2-satisfying) matrix with all
off-diagonal entries equal to a base probability $p_0$, and $\mathbf{D}$ is a
fixed ``direction'' matrix with zero diagonal, whose off-diagonal entries
create column-wise asymmetry. Off-diagonal entries of $\mathbf{P}(s)$ are
clamped to $[0.05, 0.95]$ to maintain valid probabilities. At $s = 0$, C2
holds exactly; as $s$ increases, the columns of $\mathbf{P}$ diverge and the
masking becomes increasingly informative.

\subsection{Misspecification Under C2 Violation}
\label{sec:misspec}

We now characterize what happens when the standard C1-C2-C3 model is fitted
to data generated under C2 violation.

\begin{theorem}[Pseudo-True Parameter Under C2 Misspecification]
\label{thm:misspec-C2}
Suppose data is generated under C1 and C3 with informative masking weights
$\pi_{k,c}$, but estimation is performed under C1-C2-C3 (assuming
non-informative masking). By the general theory of misspecified
MLEs~\citep{white1982misspecified, huber1967behavior}, the MLE under the
misspecified model converges in probability to a pseudo-true parameter
$\vtheta^{\dagger}$ satisfying
\begin{equation}
\label{eq:pseudo-true}
    \E_{\vtheta^*}\left[\frac{\partial \ell^{\mathrm{wrong}}}{\partial
    \theta_j}\bigg|_{\vtheta = \vtheta^{\dagger}}\right] = 0,
\end{equation}
where the expectation is taken under the true data-generating process with
parameter $\vtheta^*$. For exponential components with rates
$\vtheta = (\theta_1, \ldots, \theta_m)$, the pseudo-true parameter differs
from $\vtheta^*$ whenever the masking weights $\pi_{k,c}$ vary with $k$,
i.e., whenever C2 is violated.
\end{theorem}

\begin{proof}
The misspecified score for exponential component $j$ is
\begin{equation}
    \frac{\partial \ell_i^{\mathrm{wrong}}}{\partial \theta_j} = -s_i +
    \delta_i\frac{\ind{j \in c_i}}{\sum_{k \in c_i} \theta_k}.
\end{equation}
The true score under C1-C3 (with masking weights $\pi_{k,c}$) is
\begin{equation}
    \frac{\partial \ell_i^{\mathrm{true}}}{\partial \theta_j} = -s_i +
    \delta_i\frac{\pi_{j,c_i}\,\ind{j \in c_i}}
         {\sum_{k \in c_i} \theta_k\,\pi_{k,c_i}}.
\end{equation}
At $\vtheta^*$, the true score has expectation zero. The misspecified score
replaces the weights $\pi_{k,c}$ with uniform weights. Setting
$\E[\partial\ell^{\mathrm{wrong}}/\partial\theta_j] = 0$ at
$\vtheta^{\dagger}$ yields a system whose solution differs from $\vtheta^*$
whenever the masking weights vary with $k$.
\end{proof}

\begin{proposition}[System Hazard Robustness, exponential components]
\label{prop:hazard-robust}
Let the components be exponential, so that the system hazard is
$S = \sum_j \theta_j$, and suppose the retained system lifetimes satisfy
$T \sim \mathrm{Exp}(S^*)$ with $S^* = \sum_j \theta_j^*$ (the
masking mechanism produces retention events independent of $T$; see
\Cref{thm:exact-hazard-exp} and its corollaries). Then the
misspecified C1-C2-C3 MLE preserves the total system hazard exactly,
$\sum_j \theta_j^{\dagger} = S^*$, for any severity of C2 violation.
\end{proposition}

\begin{proof}
This is the specialization of \Cref{thm:exact-hazard-exp} to the C2
violation model, proved there via the profile-likelihood
factorization of the exponential log-likelihood into an $S$-only and a
$\phi$-only part; \Cref{cor:c2-hazard-exp} states the C2 case
explicitly. The estimand $\sum_j \theta_j$ is the system hazard
precisely because the components are exponential.
\end{proof}

\begin{remark}[Interpretation and scope]
\label{rem:interpretation}
The misspecified estimator acts as if each candidate in $c_i$ contributes
equally to the observed failure, regardless of the actual masking weights. When
a component is over-included in candidate sets (its column of $\mathbf{P}$ has
higher off-diagonal values), it ``claims credit'' for more failures than it
actually caused, inflating its estimated rate. The total system hazard is
protected because, for exponential components, the system-hazard MLE
$\hat S = n_F / \sum_i s_i$ depends on the data only through the failure
count and total exposure, neither of which is altered by how failures are
attributed across candidates. For non-exponential components $\sum_j \theta_j$
is no longer the system hazard and the exact factorization fails; total-hazard
robustness then weakens to first-order preservation in the violation severity,
which we report empirically rather than claim as an exact result.
\end{remark}

\subsection{Exact Total-Hazard Preservation for Exponential Components}
\label{sec:exact-hazard-exp}

The total-hazard robustness of \Cref{prop:hazard-robust} holds approximately
for general parametric components and is strengthened to an exact result for
the exponential case via a profile-likelihood argument. The result applies
not only to C2 misspecification but to any masking violation, including C1
and C3 violations to be analyzed in \Cref{sec:c1-sensitivity,sec:c3-sensitivity}.

\begin{theorem}[Exact Total-Hazard Preservation, Exponential Components]
\label{thm:exact-hazard-exp}
For an $m$-component series system with exponential components and true
rates $\vtheta^* = (\theta_1^*, \ldots, \theta_m^*)$, suppose data are
generated under any masking model such that retained system lifetimes
satisfy $T \sim \mathrm{Exp}(S^*)$, where $S^* = \sum_j \theta_j^*$. The
standard C1-C2-C3 MLE applied to retained observations then satisfies
\begin{equation}
\label{eq:exact-hazard-exp}
    \sum_{j=1}^{m} \theta_j^\dagger = S^*
\end{equation}
exactly, with no bias of any order in the total system hazard, regardless of
the severity of the masking violation.
\end{theorem}

\begin{proof}
Reparameterize $\theta_j = S \phi_j$ with $S = \sum_j \theta_j$ and
$\phi$ lying on the open simplex. The standard C1-C2-C3 log-likelihood for
exponential components is
\begin{equation}
    \ell^{\mathrm{wrong}}(\vtheta) =
        -S \sum_{i=1}^n s_i + \sum_{i : \delta_i = 1}
        \log\left(\sum_{k \in c_i} \theta_k\right),
\end{equation}
and substituting $\theta_k = S \phi_k$:
\begin{equation}
\label{eq:like-factored}
    \ell^{\mathrm{wrong}}(S, \phi) =
        \underbrace{- S \sum_i s_i + n_F \log S}_{\ell_S(S)}
        + \underbrace{\sum_{i : \delta_i = 1} \log W_{c_i}(\phi)}_{\ell_\phi(\phi)},
\end{equation}
where $n_F = |\{i : \delta_i = 1\}|$ and
$W_c(\phi) = \sum_{k \in c} \phi_k$. The log-likelihood is the sum of an
$S$-only part and a $\phi$-only part. The score for $S$ is therefore
$\partial \ell^{\mathrm{wrong}} / \partial S = -\sum_i s_i + n_F / S$,
independent of $\phi$, yielding the MLE
\begin{equation}
\label{eq:S-mle}
    \hat S = \frac{n_F}{\sum_i s_i}.
\end{equation}
This is the standard exponential rate MLE for type-I censored data with
$n_F$ observed failures and total observation time $\sum_i s_i$. Under the
assumed retention distribution $T \sim \mathrm{Exp}(S^*)$,
$\hat S \to S^*$ in probability, hence $\sum_j \theta_j^\dagger = S^*$.
\end{proof}

\begin{remark}[Scope]
\label{rem:exact-hazard-scope}
The result rests on two structural features. First, the exponential
log-likelihood factors additively as in \eqref{eq:like-factored}, a
consequence of the exponential hazard being constant in time. For Weibull
or other non-constant-hazard parametric components, no such factorization
holds, and total-hazard robustness generally weakens to first-order
preservation in the violation severity. Second, the marginal distribution
of retained system lifetimes must be $\mathrm{Exp}(S^*)$, which holds
whenever the masking mechanism produces retention events independent of
$T$. We verify this condition for the Bernoulli C1, C2, and C3 violation
models considered in this paper as corollaries below.
\end{remark}

\begin{corollary}[Strengthening of \Cref{prop:hazard-robust} for exponential]
\label{cor:c2-hazard-exp}
For exponential components under the Bernoulli C2 violation model with C1
holding, the misspecified C1-C2-C3 MLE preserves $S^*$ exactly, not just
approximately, for any severity of C2 violation.
\end{corollary}

\begin{proof}
Under any Bernoulli C2 violation model with C1 holding, $|C| \geq 1$
almost surely and no retention is needed. The marginal distribution of $T$
is $\mathrm{Exp}(S^*)$, so \Cref{thm:exact-hazard-exp} applies.
\end{proof}

\subsection{The Identifiability Trap}
\label{sec:identifiability}

A natural response to C2 uncertainty would be to estimate $\mathbf{P}$ jointly
with $\vtheta$ from the data. This fails.

\begin{theorem}[Non-Identifiability of $(\vtheta, \mathbf{P})$]
\label{thm:non-ident}
For exponential series systems with masked data under the Bernoulli masking
model, the joint parameter $(\vtheta, \mathbf{P})$ is not identifiable from
the observed data $(s_i, c_i, \delta_i)_{i=1}^n$. Specifically, for any
$(\vtheta^*, \mathbf{P}^*)$ with $\vtheta^* \neq \vtheta'$ there exists a
one-parameter family of $(\vtheta', \mathbf{P}')$ yielding the same
likelihood. However, the total system hazard $\sum_{j=1}^m \theta_j$ remains
identifiable from the marginal system lifetime data.
\end{theorem}

\begin{proof}
We prove the result constructively for $m = 2$ components; the general case
follows analogously. The observed-data likelihood under the Bernoulli model
depends on $(\theta_1, \theta_2, p_{12}, p_{21})$ through three functionals:
(i)~the system hazard $S = \theta_1 + \theta_2$, which determines the system
survival $R_T(t) = e^{-St}$; (ii)~the weighted hazard contribution for
candidate set $c = \{1,2\}$: $\theta_1 p_{21} + \theta_2 p_{12}$; and
(iii)~the single-component hazard contribution: $\theta_1(1 - p_{21})$ for
$c = \{1\}$, and $\theta_2(1 - p_{12})$ for $c = \{2\}$.

This gives four unknowns $(\theta_1, \theta_2, p_{12}, p_{21})$ constrained by
three likelihood-determining quantities, leaving a one-dimensional
non-identifiable family. Explicitly: given $(\theta_1^*, \theta_2^*, p_{12}^*,
p_{21}^*)$, define for any $\varepsilon$ with $|\varepsilon|$ small enough
that all parameters remain in their domains:
\begin{align}
    \theta_1' &= \theta_1^* + \varepsilon, \quad
    \theta_2' = \theta_2^* - \varepsilon, \\
    p_{21}' &= 1 - \frac{\theta_1^*(1 - p_{21}^*)}{\theta_1^* + \varepsilon},
    \quad
    p_{12}' = 1 - \frac{\theta_2^*(1 - p_{12}^*)}{\theta_2^* - \varepsilon}.
\end{align}
Then $S' = S^*$, and direct substitution verifies that all three functionals
are preserved, so $(\vtheta', \mathbf{P}')$ yields the same likelihood as
$(\vtheta^*, \mathbf{P}^*)$.

Total hazard identifiability follows from the marginal system lifetime:
$R_T(t) = e^{-St}$ depends only on $S$, which is identifiable from the
(uncensored or right-censored) system failure times alone, without reference to
the candidate sets.
\end{proof}

This extends the classical non-identifiability result for competing risks
\citep{tsiatis1975nonidentifiability, crowder2001classical} to the masked data
setting: the masking mechanism and component parameters play analogous roles to
the dependence structure and marginal distributions in the competing risks
framework~\citep{manski2003partial}.

This result is supported by both theoretical argument and simulation evidence.
Even with $n = 2000$ observations, the joint MLE of $(\vtheta, \mathbf{P})$
exhibits persistent bias with individual component rates confounded with the
off-diagonal entries of $\mathbf{P}$, while the total hazard
$\sum_j \hat\theta_j$ converges to the true value
(see~\Cref{sec:appendix-ident}).

\begin{remark}[Implications]
Since $\vtheta$ and $\mathbf{P}$ cannot be disentangled from the data, we
cannot ``just estimate'' the masking mechanism. This makes sensitivity
analysis the appropriate tool: fit the standard C1-C2-C3 model, then assess
how conclusions change under plausible perturbations of the masking structure
via different $\mathbf{P}$ matrices. If the substantive conclusions are stable,
C2 violations are not a concern for the application at hand.
\end{remark}

\subsection{Sensitivity to C1 Violations}
\label{sec:c1-sensitivity}

We now extend the framework to violations of C1, the assumption that the
failed component is in the candidate set. The construction is parallel to
the C2 case of \Cref{sec:bernoulli,sec:misspec}: we introduce a calibrated
scalar severity parameter, derive the misspecified score equation, and
expand the pseudo-true parameter to first order.

\paragraph{The Bernoulli C1 violation model.}
Within the Bernoulli masking model of \Cref{def:bernoulli}, a C1 violation
is induced by setting
\begin{equation}
\label{eq:c1-violation}
    p_k(k) = 1 - \alpha_{C_1}, \qquad
    p_j(k) = p_0 \ \text{for } j \neq k.
\end{equation}
The off-diagonal probabilities are constant in $k$, so C2 continues to
hold along the C1 sweep, isolating the effect of C1 violation. The
parameter $\alpha_{C_1} \in [0, 1/2]$ measures the probability that the
true cause is excluded from the candidate set, with $\alpha_{C_1} = 0$
recovering the standard model and $\alpha_{C_1} = 1/2$ marking the boundary
beyond which the candidate set is anti-informative (the true cause is
more likely excluded than included).

Under \eqref{eq:c1-violation} the candidate set may be empty with
probability $\alpha_{C_1} (1 - p_0)^{m-1}$ conditional on a failure. The
standard C1-C2-C3 likelihood is undefined on empty candidate sets, so we
condition on $|C| \geq 1$ throughout the analysis. This matches the
practitioner workflow: failures with no diagnostic information are
excluded from the parametric MLE.

\begin{theorem}[Pseudo-True Parameter Under C1 Misspecification, $m=2$]
\label{thm:misspec-C1}
For a two-component exponential series system with rates
$\vtheta^* = (\theta_1^*, \theta_2^*)$ and data generated under the
Bernoulli C1 violation model \eqref{eq:c1-violation}, conditioning on
$|C| \geq 1$, the misspecified C1-C2-C3 MLE converges to a pseudo-true
parameter $\vtheta^\dagger(\alpha_{C_1})$ admitting the first-order
expansion
\begin{equation}
\label{eq:c1-expansion}
    \theta_j^\dagger = \theta_j^* +
        \alpha_{C_1}\, \Delta_j^{C_1}(\vtheta^*; p_0) +
        O(\alpha_{C_1}^2),
\end{equation}
with
\begin{equation}
\label{eq:c1-delta}
    \Delta_1^{C_1} = \frac{p_0\,(\theta_2^* - \theta_1^*)}{1 - p_0},
    \qquad
    \Delta_2^{C_1} = -\Delta_1^{C_1} =
        \frac{p_0\,(\theta_1^* - \theta_2^*)}{1 - p_0}.
\end{equation}
\end{theorem}

\begin{proof}
The misspecified score for exponential component $j$ is
$\partial \ell_i^{\mathrm{wrong}} / \partial \theta_j = -s_i + \delta_i \,
\ind{j \in c_i} / \sum_{k \in c_i} \theta_k$. Setting
$\E_{\vtheta^*}[\partial \ell^{\mathrm{wrong}} / \partial \theta_j] = 0$ at
$\vtheta = \vtheta^\dagger$, and computing the expectation under the
Bernoulli C1 violation model with $|C| \geq 1$ and no censoring, the score
equation for $\theta_1^\dagger$ reduces to (after dividing by $S^*$ and
multiplying by the conditional-probability normalizer
$Z = 1 - \alpha_{C_1}(1-p_0)$):
\begin{equation}
\label{eq:c1-score}
    Z =
    \frac{\theta_1^*(1-\alpha_{C_1})(1-p_0) + \theta_2^* \alpha_{C_1} p_0}
         {\theta_1^\dagger}
    + \frac{p_0 (1-\alpha_{C_1}) S^*}{\theta_1^\dagger + \theta_2^\dagger},
\end{equation}
with the analogous equation for $\theta_2^\dagger$ obtained by swapping
indices. Writing $\theta_j^\dagger = \theta_j^* + \alpha_{C_1} \Delta_j +
O(\alpha_{C_1}^2)$ and expanding both sides of \eqref{eq:c1-score} to
first order in $\alpha_{C_1}$, the zeroth-order terms vanish identically.
The first-order coefficients yield the linear system
\begin{align}
\label{eq:c1-linsys-1}
    (1-p_0)\, \frac{\Delta_1}{\theta_1^*}
        + p_0\, \frac{\Delta_1 + \Delta_2}{S^*}
        &= p_0\, \frac{\theta_2^* - \theta_1^*}{\theta_1^*}, \\
\label{eq:c1-linsys-2}
    (1-p_0)\, \frac{\Delta_2}{\theta_2^*}
        + p_0\, \frac{\Delta_1 + \Delta_2}{S^*}
        &= p_0\, \frac{\theta_1^* - \theta_2^*}{\theta_2^*}.
\end{align}
Setting $a = \Delta_1 / \theta_1^*$, $b = \Delta_2 / \theta_2^*$, and
$r = \theta_2^* / \theta_1^*$, the system \eqref{eq:c1-linsys-1},
\eqref{eq:c1-linsys-2} reduces to two linear equations in $(a, b)$.
Combining them yields $a + br = 0$, i.e., $\Delta_1 + \Delta_2 = 0$.
Substituting back gives
$a = p_0 (r-1) / (1 - p_0)$, hence
$\Delta_1 = \theta_1^* \cdot p_0 (r-1) / (1-p_0) = p_0 (\theta_2^* -
\theta_1^*) / (1-p_0)$, and similarly $\Delta_2 = -\Delta_1$.
\end{proof}

\begin{corollary}[Total Hazard Preservation Under C1 Violation, Exponential, General $m$]
\label{cor:c1-hazard-general}
For exponential series systems of any size $m \geq 2$ under the Bernoulli
C1 violation model \eqref{eq:c1-violation} with retention on
$|C| \geq 1$, the misspecified MLE preserves the total system hazard
exactly: $\sum_j \theta_j^\dagger = S^*$ for all
$\alpha_{C_1} \in [0, 1/2]$. In particular, the $m=2$ first-order
coefficients in \eqref{eq:c1-delta} satisfy $\Delta_1^{C_1} +
\Delta_2^{C_1} = 0$, and this preservation holds at all orders for
general $m$.
\end{corollary}

\begin{proof}
By \Cref{thm:exact-hazard-exp}, it suffices to verify that retained
system lifetimes have the marginal distribution $\mathrm{Exp}(S^*)$.
Under \eqref{eq:c1-violation},
$\Prob\{|C| \geq 1 \mid K = k, T = t\} = 1 - \alpha_{C_1}(1-p_0)^{m-1}$
is independent of both $k$ and $t$. The retention event $\{|C| \geq 1\}$
is therefore independent of $T$, and the conditional distribution
$T \mid |C| \geq 1$ equals the marginal $T \sim \mathrm{Exp}(S^*)$.
\end{proof}

\begin{remark}[Interpretation]
\label{rem:c1-interp}
When the candidate set excludes the true cause, the misspecified MLE
attributes the failure to whichever non-failed component appears in the
candidate set. Equation \eqref{eq:c1-delta} shows that the first-order bias
is proportional to the rate gap $\theta_2^* - \theta_1^*$: the
higher-rate component absorbs failures it did not cause from the
lower-rate component, biasing its estimate upward by the same magnitude
that the lower-rate component is biased downward. The total system hazard
is preserved because the system lifetime $T \sim \mathrm{Exp}(S^*)$
identifies $S^*$ directly from time data, independent of how individual
failures are attributed. The amplification factor $p_0 / (1 - p_0)$
reflects how much the surviving candidate set covers non-failed
components; it diverges as $p_0 \to 1$, where individual rates become
non-identifiable even under C1.
\end{remark}

The proof generalizes to $m \geq 3$ components with the same
total-hazard preservation result. The structure of the linear system
(score Jacobian on the left, $\alpha_{C_k}$-derivative of the score on
the right) is identical to \eqref{eq:c3-linsys} and is implemented
numerically for any $m$ in the \texttt{mdrelax} R package
(\Cref{app:software}).

\subsection{Sensitivity to C3 Violations}
\label{sec:c3-sensitivity}

C3 parameterizes whether the masking probability depends on the component
reliability parameters $\vtheta$. We adopt the power-weight model
\citep{Mukhopadhyay2006}:
\begin{equation}
\label{eq:c3-violation}
    p_j(k; \vtheta) = p_0 \cdot
        \frac{\theta_j^{\alpha_{C_3}}}
             {\max_{l} \theta_l^{\alpha_{C_3}}}
    \quad \text{for } j \neq k,
\end{equation}
with $p_k(k) = 1$ (C1 holds). The normalizer
$\max_l \theta_l^{\alpha_{C_3}}$ is independent of the failed
component $k$, so the inclusion weight of component $j$ is the same
regardless of which component failed: C2 holds and the violation is
purely C3 (dependence of the masking probability on $\vtheta$). This
is the model implemented in \texttt{wei\_inclusion\_probs}
(\texttt{R/wei\_series\_c1\_c2.R}), where the weights are normalized
by their global maximum. The scalar
$\alpha_{C_3} \in [0, \alpha_{C_3}^{\max}]$ is the calibrated severity:
$\alpha_{C_3} = 0$ recovers constant masking $p_j(k) = p_0$ (C3 holds),
while increasing $\alpha_{C_3}$ makes high-rate components more likely to
appear in candidate sets, modeling a diagnostic procedure that weights
candidates by their estimated reliability ranking.

\begin{theorem}[Pseudo-True Parameter Under C3 Misspecification]
\label{thm:misspec-C3}
For exponential series systems with data generated under the power-weight
C3 violation \eqref{eq:c3-violation}, the misspecified C1-C2-C3 MLE
converges to a pseudo-true parameter $\vtheta^\dagger(\alpha_{C_3})$
admitting the first-order expansion
\begin{equation}
\label{eq:c3-expansion}
    \theta_j^\dagger = \theta_j^* +
        \alpha_{C_3}\, \Delta_j^{C_3}(\vtheta^*; p_0) +
        O(\alpha_{C_3}^2).
\end{equation}
The first-order coefficients satisfy
\begin{enumerate}
    \item \textbf{Total preservation, general $m$:}
        $\sum_{j=1}^m \Delta_j^{C_3}(\vtheta^*; p_0) = 0$, and the
        all-orders identity $\sum_j \theta_j^\dagger(\alpha_{C_3}) = S^*$
        holds for every $\alpha_{C_3}$ (by \Cref{thm:exact-hazard-exp},
        since the power-weight model makes $|C| \geq 1$ deterministic so
        the retention condition is satisfied trivially).
    \item \textbf{Linear system for general $m$:} the vector
        $\Delta^{C_3}(\vtheta^*; p_0) \in \mathbb{R}^m$ is the unique
        solution (on the rate simplex) of the score-Jacobian system
        \begin{equation}
        \label{eq:c3-linsys}
            \mathbf{J}(\vtheta^*)\, \Delta^{C_3}(\vtheta^*; p_0)
                \,=\, \mathbf{r}(\vtheta^*; p_0),
        \end{equation}
        where $\mathbf{J}(\vtheta^*)$ is the Jacobian of the misspecified
        score in $\vtheta$ at $\vtheta^*$ (the standard expected
        information matrix evaluated under $\alpha_{C_3} = 0$) and
        $\mathbf{r}(\vtheta^*; p_0)$ is the partial derivative of the
        misspecified score in $\alpha_{C_3}$ at $\alpha_{C_3} = 0$
        evaluated under the power-weight perturbation. Both matrices are
        rational functions of $(\vtheta^*, p_0)$ obtained by
        differentiating the misspecified score in
        \Cref{thm:misspec-C2}. The total-preservation identity in item~1
        forces $\mathbf{1}^\top \Delta^{C_3} = 0$, so the system has rank
        $m - 1$ on the simplex; the rank deficiency is exactly the
        directional ambiguity that the masking-mechanism non-identifiability
        of \Cref{thm:non-ident} predicts.
\end{enumerate}
\end{theorem}

The simulation results in \Cref{sec:simulations} demonstrate the
qualitative behavior of $\Delta_j^{C_3}$: rates of high-failure-rate
components are inflated further (positive bias amplification), while
low-rate components are underestimated. The component-level redistribution
respects the total-hazard constraint for exponential components. For
Weibull components, where the factorization \eqref{eq:like-factored} does
not apply, total-hazard preservation under C3 is no longer automatic and
is studied empirically in \Cref{sec:simulations}.

\subsection{Local Sensitivity Indices and the Robustness Hierarchy}
\label{sec:hierarchy}

The first-order coefficients $\Delta_j^{C_k}$ from
\Cref{thm:misspec-C2}, \Cref{thm:misspec-C1}, and \Cref{thm:misspec-C3}
specialize the local-sensitivity construction of
\citet{TroxelMaHeitjan2004} and \citet{ZhangHeitjan2006} to candidate-set
coarsening. We collect them into a single index and use it to classify
estimands by robustness.

\begin{definition}[Local Sensitivity Index for Candidate-Set Coarsening]
\label{def:isni}
For a functional $\psi(\vtheta)$ of the parameter vector and a violation
type $C_k \in \{C_1, C_2, C_3\}$, the local sensitivity index is
\begin{equation}
\label{eq:isni}
    \mathrm{ISNI}_{C_k}[\psi]
    \,\coloneqq\,
    \left.
        \frac{\partial}{\partial \alpha_{C_k}}
        \psi\bigl(\vtheta^\dagger(\alpha_{C_k})\bigr)
    \right|_{\alpha_{C_k} = 0}
    = \nabla \psi(\vtheta^*)^{\!\top}\, \Delta^{C_k}(\vtheta^*).
\end{equation}
\end{definition}

\paragraph{Robustness intervals.}
The local sensitivity index converts an unverifiable masking deviation
$\alpha_{C_k}$ into a bound on the asymptotic bias of $\psi$ under the
standard MLE. We package that bound as an interval covering the
pseudo-true value $\psi(\vtheta^\dagger(\alpha_{C_k}))$ across a
practitioner-chosen plausibility set for $\alpha_{C_k}$.

\begin{definition}[First-Order Robustness Interval]
\label{def:robustness-interval}
Fix a functional $\psi \colon \Theta \to \mathbb{R}$ that is $C^2$ in
a neighborhood of the standard-model fit $\vtheta^*$, a violation type
$C_k$, and a plausibility set $\mathcal{A}_k \subset [0,
\alpha_{C_k}^{\max}]$. The first-order robustness interval for $\psi$
under $C_k$ is
\begin{equation}
\label{eq:rob-int}
    \mathrm{RI}_{C_k}^{(1)}(\psi; \mathcal{A}_k)
    \,\coloneqq\,
    \Bigl[
        \widehat\psi
        \,-\, |\mathrm{ISNI}_{C_k}[\psi]|\, \sup_{\alpha \in \mathcal{A}_k} \alpha,
        \ \widehat\psi
        \,+\, |\mathrm{ISNI}_{C_k}[\psi]|\, \sup_{\alpha \in \mathcal{A}_k} \alpha
    \Bigr],
\end{equation}
where $\widehat\psi = \psi(\widehat\vtheta)$ is the plug-in estimate
from the standard MLE.
\end{definition}

\begin{proposition}[Asymptotic Coverage of $\psi(\vtheta^\dagger)$]
\label{prop:robustness-coverage}
Assume the regularity conditions of \Cref{thm:misspec-C2,thm:misspec-C1,thm:misspec-C3}:
the standard-model fit $\widehat\vtheta$ is $\sqrt{n}$-consistent for
$\vtheta^*$ when $\alpha_{C_k} = 0$; the pseudo-true map
$\alpha_{C_k} \mapsto \vtheta^\dagger(\alpha_{C_k})$ is $C^2$ on a
neighborhood of $0$ with a uniform bound $M_k$ on the operator norm of
its Hessian and a uniform bound $D_k$ on the norm of its first
derivative $\dot\vtheta = \mathrm{d}\vtheta^\dagger/\mathrm{d}\alpha_{C_k}$
on $\mathcal{A}_k$; and $\psi$ is $C^2$ with $\|\nabla\psi\|$ bounded
by $L_\psi$ and $\|\nabla^2\psi\|$ bounded by $B_\psi$ on a
neighborhood of $\vtheta^*$. Then for every $\alpha_{C_k} \in
\mathcal{A}_k$ the pseudo-true value admits the expansion
\begin{equation}
\label{eq:rob-sandwich}
    \psi\bigl(\vtheta^\dagger(\alpha_{C_k})\bigr)
    \,=\, \psi(\vtheta^*)
    \,+\, \alpha_{C_k}\, \mathrm{ISNI}_{C_k}[\psi]
    \,+\, R_k(\alpha_{C_k}),
    \qquad
    |R_k(\alpha_{C_k})| \leq \tfrac{1}{2}\bigl(L_\psi M_k + B_\psi D_k^2\bigr)\, \alpha_{C_k}^2,
\end{equation}
with $L_\psi = \sup_{\vtheta \in \mathcal{N}(\vtheta^*)} \|\nabla
\psi(\vtheta)\|$ and $B_\psi = \sup_{\vtheta \in
\mathcal{N}(\vtheta^*)} \|\nabla^2\psi(\vtheta)\|$. Consequently,
$\mathrm{RI}_{C_k}^{(1)}(\psi; \mathcal{A}_k)$ contains
$\psi(\vtheta^\dagger(\alpha_{C_k}))$ for every $\alpha_{C_k} \in
\mathcal{A}_k$ up to a remainder that is $O(\sup \mathcal{A}_k^2)$.
\end{proposition}

\begin{proof}
Write $f(\alpha) = \psi(\vtheta^\dagger(\alpha))$. By the chain rule,
$f'(\alpha) = \nabla\psi(\vtheta^\dagger)^\top \dot\vtheta$ and
\begin{equation*}
    f''(\alpha) = \dot\vtheta^\top \nabla^2\psi(\vtheta^\dagger)\,\dot\vtheta
    \;+\; \nabla\psi(\vtheta^\dagger)^\top \ddot\vtheta,
\end{equation*}
so the second-order term carries \emph{both} the curvature of the
pseudo-true map ($\ddot\vtheta$, bounded by $M_k$) and the curvature
of the estimand ($\nabla^2\psi$ contracted with $\dot\vtheta^{\otimes
2}$, bounded by $B_\psi D_k^2$). The linear term is
$f'(0) = \mathrm{ISNI}_{C_k}[\psi]$. The Lagrange remainder of the
first-order Taylor expansion is $\tfrac12 f''(\xi)\alpha^2$ for some
$\xi$, and $|f''(\xi)| \leq L_\psi M_k + B_\psi D_k^2$ uniformly on
$\mathcal{A}_k$, giving \eqref{eq:rob-sandwich}. The estimand-curvature
term $B_\psi D_k^2$ is nonzero for every nonlinear $\psi$ (MTTF,
reliability, hazard-at-$t$) and vanishes only for affine $\psi$. The
interval in \eqref{eq:rob-int} contains the linear term for every
$\alpha_{C_k} \in \mathcal{A}_k$ by construction.
\end{proof}

\begin{remark}[How to use the interval]
\label{rem:robustness-interval-usage}
The practitioner picks $\mathcal{A}_k = [0, \alpha_{C_k}^\star]$ where
$\alpha_{C_k}^\star$ is the largest violation deemed plausible for the
diagnostic process at hand: the default scale is calibrated so that
$\alpha_{C_k}^\star = \alpha_{C_k}^{\max}$ is the boundary of the
adversarial regime. The ISNI value
$\mathrm{ISNI}_{C_k}[\psi]$ is obtained in closed form from
$\nabla \psi(\widehat\vtheta)$ and the $\Delta^{C_k}(\widehat\vtheta)$
expression from
\Cref{thm:misspec-C2,thm:misspec-C1,thm:misspec-C3}. The interval is
deliberately conservative: it bounds the asymptotic pseudo-true bias and
does not absorb Monte Carlo uncertainty in $\widehat\vtheta$, which can
be added by widening the interval by the usual Wald half-width
$1.96\, \widehat{\mathrm{se}}(\widehat\psi)$.
\end{remark}

\begin{remark}[When the first-order interval collapses]
\label{rem:rob-int-collapse}
For estimands that the hierarchy classifies as exactly preserved
(items 1 and 2 of \Cref{cor:hierarchy}) the ISNI is zero,
$\mathrm{RI}_{C_k}^{(1)}$ collapses to the point estimate, and
\Cref{prop:robustness-coverage} becomes the assertion that the
pseudo-true value coincides with the standard-model fit at every order.
For estimands that are first-order robust but not exactly preserved
(item 3) the interval has zero linear half-width and only the
second-order remainder $R_k$ matters; in practice this remainder is
absorbed into the Monte Carlo or sandwich variance. For estimands that
are not first-order robust (item 4) the interval width scales linearly
with $\sup \mathcal{A}_k$, and is the operative bound for
practitioner-facing component-level claims.
\end{remark}

\begin{corollary}[Robustness Hierarchy]
\label{cor:hierarchy}
For masked series systems under the standard C1-C2-C3 MLE:
\begin{enumerate}
    \item \textbf{Exponential components, total system hazard
        $S(\vtheta) = \sum_j \theta_j$:} preserved exactly under any
        violation of C1, C2, or C3
        (\Cref{thm:exact-hazard-exp}, \Cref{cor:c2-hazard-exp},
        \Cref{cor:c1-hazard-general}), giving $\mathrm{ISNI}_{C_k}[S] = 0$
        for all $k$.
    \item \textbf{Exponential components, mean time to system failure
        $1/S(\vtheta)$:} preserved exactly under any of the three
        violations (immediate from item 1).
    \item \textbf{Weibull components, system-level estimands (system
        hazard at fixed $t$, MTTF):} first-order robust under C2 and C3
        violations ($\mathrm{ISNI}_{C_2}[S] = \mathrm{ISNI}_{C_3}[S] = 0$
        at the standard-model fit). Under C1, the first-order coefficient
        is nonzero but the absolute bias remains bounded across the
        practical violation range $\alpha_{C_1} \in [0, 0.5]$, with
        empirical magnitude below $4\%$ in the 5-component sweep of
        \Cref{sec:simulations}.
    \item \textbf{Individual component parameters} (rates $\theta_j$ for
        exponential, shape-scale pairs for Weibull): not first-order
        robust under any of the three violations. Scale parameters of
        low-rate components exhibit the most catastrophic degradation.
\end{enumerate}
\end{corollary}

The asymmetry in item 3 has a structural origin. The exact-preservation
proof for exponential (\Cref{thm:exact-hazard-exp}) relies on the
log-likelihood factorization $\ell(S, \phi) = \ell_S(S) + \ell_\phi(\phi)$,
which decouples system-level inference from component-level inference. For
Weibull components the analogous factorization fails: the time-portion of
the likelihood involves the individual shape-scale pairs that also appear
in the candidate-set portion. Under C2 and C3 violations, the
candidate-set bias remains balanced enough that it does not propagate to
the time portion at first order. Under C1 violation, the candidate set
sometimes omits the true cause entirely, producing more aggressive
candidate-set bias that does propagate to the time-portion estimates at
first order. The bias remains bounded because the system lifetime
$T \sim \mathrm{Weibull}_m^{\mathrm{series}}(\vtheta^*)$ still satisfies
the marginal-distribution preservation that powers the exponential
argument; only the parametric mapping back to $S(t)$ is biased.

The hierarchy formalizes the empirical pattern observed in
\Cref{sec:simulations}. The practical takeaway is sharp: practitioners
reporting system-level reliability estimates from exponential models can
ignore concerns about masking-condition violations entirely. For
Weibull-component systems, system-level estimates are first-order
protected against C2 and C3 violations and admit a small, bounded bias
under C1. Component-level estimates are sensitive to all three violations
regardless of the component family, and require sensitivity analysis
before being reported. We discuss the practical consequences in
\Cref{sec:discussion}.
